\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-08-30 notes/2loopNLSM.tex
small edits

\subsubsection{Jacobi relations}
\label{sec:jac-rels}

@@

\subsubsection{Graph symmetries}
\label{sec:syms}

@@

\subsubsection{Unitarity}
\label{sec:unitarity}

@@

\section{Isolating the cubic sector}
Here we'll describe the why constructing the full Yang-Mills kinematic algebra (outside self-dual sector) is generically hard, and why the same reason makes NLSM beyond one-loop hard as well. It has to do with out the vector state sum mixes with the four-point contact. In terms of our previously defined Feynman rules, the three-point YM vertex takes the form:
\be
\cubic{hgrey0}{gluon}{}{gluon}{}{gluon}{}  =  \cubic{hgrey0}{fermion2}{}{}{nhpRed}{}{nhpRed} + \cubic{hgrey0}{}{nhpRed}{fermion2}{}{}{nhpRed} + \Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{} +\text{h.c.}
\ee
The state-sum between two three-point gluon vertex yields the following expansion in terms of YZ-theory vertices:
\be
 \AcubicB{hgrey0}{gluon}{}{gluon}{}{gluon}{}{gluon}{} =  \AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{fermion2}{} + 5 \text{ perms} + \AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{}{nhpRed}
\ee
All the $(\epsilon\epsilon)^1$ diagram permutations are isolated to the cubic sector and together manifest color-kinematics duality. However, the exception of the last diagram which introduces a $(\epsilon\epsilon)^2$ factor, must conspire with the Yang-Mills contact to preserve the duality off-shell

\be
\AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{}{nhpRed} = (s_{13}-s_{12}) \epsilon_{(12)} \epsilon_{(34)}
\ee
One was to resolve this is to introduce higher spin modes in order to accomodate the internal index contractions needed to recover the s-channal contract:
\be
\AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{Bfield}{nhpRed}{}{nhpRed} = s_{12 }(\epsilon_{(13)} \epsilon_{(24)} - \epsilon_{(14)} \epsilon_{(23)})
\ee
While introducing succesively higher spiin states to the NMHV lagrangian constructed in \citepls{} could in principle work at tree level, we will see that it is not consiisntent with what we find at general loop order. We will comment on this in \sect{}. For now, we will study how to construct one-loop integrands consistent with the color-kinematics by isolating the cubic sector of the state sum above. This is easily understood in terms of a polarization decomposition introduced by one of the authors \citepls{} that decomposes vector amplitudes in terms of building blocks that preserve partial amplitude relations. 
\subsection{Block Decomposition of Yang-Mills}
Describe how different orders in $\varepsilon_i\varepsilon_j$ in pure Yang-Mills captures different cubic and contact sectors of the theory. All the $D$-dimensional Lagnrangians above can be understood as selecting specific on shell states to isolate the cubic sector of Yang-Mills at a given loop order. Start with reducible amplitude block decomp from Ref. \citepls{}:
\begin{eBox}
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
\end{eBox}
By a simple mass-dimensioin argument, we know the tree-level cubic sector is simply:
\be
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
and likewise at one-loop:
\be
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
We can select out both of these manifestly cubic sectors by selecting only $\epsilon^+\epsilon^+$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level OR what $J$-theory and $YZ$-theory select with their on-shell states. Indeed, the $D$-dimensional description of $J$ theory is just the following blocks in the RABD expansion of Yang-Mills:
\be
A(...,\bar{J}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
\ee
and similarly so at one-loop:
\be
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}
\ee
To recover NLSM at one-loop, we need to extract out the $D$-dependent part that corresponds to an internal $YY$-loop from extra-dimensional scalars:
\begin{eBox}
\be
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\epsilon\rightarrow k}
\ee
\end{eBox}
Why must we take the derivative with respect to $D$? After all, $J$-theory and $YZ$-theory both have well defined propagators for internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we noted in the previous section, all the $J$ and $Z$ states must be on-shell in order to produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at one-loop will not produce NLSM amplitudes. However, the internal $YY$-loop is a valid forward limit, since as constructed YZ theory is valid for off-shell Y-particles, and thus they serve as a valid forward limit. We will construct these one-loop integrands below in the next section. 
\subsection{One-loop SDYM Integrands}
\be
\simplebox = g_{\text{YM}}^4 X(k_1,l_1)X(k_2,l_2)X(k_3,l_3)X(k_4,l_4)
\ee
\subsection{One-loop NLSM Integrands}

Note from James:  I've explicitly checked the nice form of the 1-loop numerators at 4pt, 6pt, and 8pt.
\be
\simplebox = (2 f_\pi^{-1})^4 (l_1 k_1)(l_2 k_2)(l_3 k_3)(l_4 k_4)
\ee
where we define sequential loop momenta on the $n$-gon as $l_{i+1}= l_i+k_i$. Given this, the $n$-gon permits a equivalent expression in terms of inverse propagators:
\be
\simplebox =f_\pi^{-4} [12][23][34][41]
\ee
Where we have define the antisymmetric kinematic variable, $[LR] = l_i^L - l_R^2$
\subsection{Two-loop Obstruction}

\section{Bootstrapping two-loop...}

@@

\subsection{Double-copy -- Alex }
\ace{I'm punting a lot of discussion to Ref \cite{Carrasco:2023qgz}.
  Do we think that is a problem?}

\begin{figure}
  \begin{center}
    \begin{subfigure}{0.45\textwidth}
      \begin{center}
        \PhysicalCutOne{}{}{}{}
      \end{center}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
      \begin{center}
        \PhysicalCutTwo{}{}{}{}
      \end{center}
    \end{subfigure}
  \end{center}
  \caption{The two physical cut topologies shared by sGal, BI, and DBIVA}
  \label{fig:emu}
\end{figure}

@@

\subsection{Yang-Mills revisited -- Alex}
Given the close ties between pion and gluon scattering for trees and
at one loop, the success of a globally color-dual solution for pions
at two loops warrants a reexamination of the two-loop pure Yang--Mills
problem.  As discussed above in \cref{sec:bootstrap}, BDN failed at
finding an off-shell color-dual representation satisfying 7
assumptions, and were forced to develop an on-shell notion of
color-kinematics duality \cite{Bern:2015ooa}.  Thus we must decide
which assumption(s) to break.  From both a motivation and
implementation standpoint, the simplest assumption to break is the
``power-counting'' statement, which we analyze first.

@@

      \ver{c1}{0:\cl};
      \ver{c2}{\cang:\cl};
      \ver{c3}{2*\cang:\cl};
      \ver{c4}{3*\cang:\cl};
      \ver{ci}{$(i5)!0.5!(i1)$};
      \path let \p1 =(ci) in coordinate (c5) at (\x1,0.2){};
      \draw[ultra thick] (e1)--(i1);
      \draw[ultra thick] (e2)--(i2);
      \draw[ultra thick] (e3)--(i3);
      \draw[ultra thick] (e4)--(i4);
      \draw[ultra thick] (i5)--(i1)--(i4)--(i3)--(i2)--(i6);
      \draw[ultra thick] (i6) to[out=60,in=180-60] (i5);
      \draw[ultra thick] (i6) to[out=-60,in=180+60] (i5);
      \draw[dashed,ultra thick,blue] (0,0) -- (c1);
      \draw[dashed,ultra thick,blue] (0,0) -- (c2);
      \draw[dashed,ultra thick,blue] (0,0) -- (c3);
      \draw[dashed,ultra thick,blue] (0,0) -- (c4);
      \draw[dashed,ultra thick,blue] (c5) -- ++(90:0.7*\cl);
      \node[fill=white,circle] (c) at (0,0) {};

@@

\ace{\sout{I need to check this.  For now it is a standing
    hypothesis:} Checked.  The failure works as stated.} In fact, we can make the
failure extremely precise.  Start with the diagrams shown in the
Jacobi relation from \cref{eq:TwoLoopJacobiOne}: the double-box,
crossed-box, and penta-triangle.  For each of the three diagrams,
write down the most generic parity-even local ansatz involving four
polarizations and six momenta (any combination of external or loop).
Each ansatz will be 10,010 terms initially.  Next impose the symmetry
constraints on each of the three diagrams, leaving 2761 free
parameters on the double-box, 2576 free parameters on the crossed-box,
and 5040 free paramters on the penta-triangle.  Then impose the
``kissing triangle'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only recieves contributions from the
double-box
\begin{equation}
  \KissingTriangles
  =
  \frac{1}{(\textcolor{ucp-color}{\ell_1})^2} \
  n \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{(\textcolor{ucp-color}{\ell_2})^2} \
  n\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied:
\begin{equation}
  \KissingTriangles
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang--Mills representation: locality,
symmetry, and the ``kissing triangle'' cut are incompatible with the
kinematic Jacobi relations.

@@

\section{... and Beyond}
This could be a section about our expectations for all-loop construction requiring non-local color-dual integrands. Lay out contrafactual proof, i,e. define the following statement:
\be
 Q_L=\textit{``the $L$-loop numerator is polynomial and color-dual"}. 
\ee
Then it is clear that if the kinematic numerator is polynomial and color-dual, evaluating it in kinematic regions where the internal legs are on-shell won't disrupt the kinematic algrebraic relations, and it must still be polynomial. Thus we find:
\be
Q_L \Rightarrow Q_{L-1}
\ee
Or equivalently,
\be
\text{not }Q_{L-1}\Rightarrow \text{not }Q_L
\ee
We have found that the two-loop integrand for Yang-Mills cannot be constructed from a color-dual polynomial kinematic numerator. Thus, if a color-dual representation exists at all-loop order for Yang-Mills it must be a rational function of kinematics.

@@

\section{Spinor-Helicity and Conventions}

Before proceeding, we note a strange property of the $n$-gon numerator for the pions. In this form, Jacobi relations produce \textit{non-vanishing} values for bubbles on external legs (BELs). However, it is easy to see that the contribution integrates to zero after applying IBP relations/tensor reduction on the bubble. The BEL diagram can be reconstructed from the $n$-gon as follows:
\be
N^{\text{BEL}}_{1|2,34} = N^{\text{NLSM}}_{(1234,l)}-N^{\text{NLSM}}_{(1243,l)}-N^{\text{NLSM}}_{(1342,l)}+N^{\text{NLSM}}_{(1432,l)}
\ee
where we define the loop momentum to be in between the left most and right most leg on the box. Plugging in particular values for $l_i$, we obtain the following expression for the BEL
\be
N^{\text{BEL}}_{1|2,34} = s_{12} (l+k_1)^2 l^{\mu} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} 
\ee
Notice there is an overall factor that cancels one of the propagators. Plugging this in, produces an integral of the following form
\be
\mathcal{I}^{\text{BEL}}_{1|2,34} = s_{12} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} \int \frac{d^D l}{i\pi^{D/2}} \frac{l^\mu }{l^2-\mu^2} \sim   s_{12}(s_{13}-s_{14}) (\mu^2)^{D/2}
\ee
where we have introduced a mass regulator that will be proportional the the on-shell momentum inside the BEL, $\mu^2 \equiv k_1^2$. Thus, in large enough dimension, this integral suppresses the $\mu^{-2}$ divergence appearing in the denominator of the BEL diagram.

\section{Review of known theories}

@@

\subsubsection{Self-Dual Yang-Mills} The first example in the literature \citepls{} of color-dual Feynman rules is that of Yang-Mills theory in the self-dual sector (SDYM):

@@

\subsubsection{YZ-Theory}
SDYM makes explicit reference to the spacetime dimension and Zakharov-Mikhailov theory, to be discussed in \cref{sec:ZMTheory}, does as well.
This makes both theories poorly suited for constructing dimensionally-regulated loop-level amplitudes.
On the other hand, YZ-theory is an honest $D$-dimensional theory that captures the classical physics of the nonlinear sigma model (NLSM) while manifest the duality between color and kinematics \cite{Cheung:2016prv}.

@@

\subsubsection{J-theory} There is an alternative color-dual formulation of NLSM that the theory manifestly corresponds to NLSM at least at tree level \cite{Cheung:2021zvb}.
For related formulations of NLSM that do not treat color-kinematics duality see \cite{Freedman:1980us, Slavnov:1971mz}.
The construction begins with a first order formulation of NLSM in terms of chiral currents $J^\mu$.
First order formulation of NLSM in terms of chiral current $J^\mu$.
Two equations.
Two constraints, the field strength should vanish
\be
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0
\ee
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge
\be
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0
\ee
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\be
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\ee
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian
\begin{eBox}
\be
\mathcal{L}^{\text{J}} =\bar{J}^\mu \Box J_\mu +g f^{abc} \bar{J}^\mu J^\nu \partial_\nu J^\mu .
\ee
\end{eBox}
By calculating an off-shell 4pt correlation function it is possible to check that this Lagrangian automatically satisfies color-kinematics duality.
Kinematic algebra is diff algebra (as usual).
How to extract physical states... (longitudinal J, weird J bar, restores broken Bose).

@@

\subsubsection{Zakharov-Mikhailov theory}
\label{sec:ZMTheory}
While J-theory manifests color-kinematics duality, it only does so at tree level because J-theory breaks manifest Bose symmetry.
Manifest color-kinematics duality can be achieved at all loop orders by restricting J-theory to two spacetime dimensions where the chiral current can be dualized to a scalar $J^\mu = \epsilon^{\mu\nu} \partial_\nu \varphi$, which trivially enforces Lorenz gauge.
After plugging $\varphi$ into \cref{eq:JTheoryEOM} and rearranging, another copy of $\varphi$ (rather than some new field) can be integrated in to obtain the Zakharov-Mikhailov (ZM) Lagrangian
\begin{eBox}
\be
\label{eq:ZMLagrangian}
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2 + g f^{abc} \varphi^a \epsilon^{\mu\nu}(\partial_\mu \varphi^b)( \partial_\nu \varphi^c) =  \text{tr}\left[\frac{1}{2}(\partial \varphi)^2-i g \varphi [\partial_t \varphi, \partial_z \varphi ]\right] .
\ee
\end{eBox}
This procedure makes it clear that ZM and NLSM encode the same classical physics, that is, their equations of motion are dual to each other.
ZM theory can also be tenuously motivated from SDYM theory.
The power counting of the interaction term in the SDYM equation of motion \cref{SDEOM} makes it clear that the theory is critical in 2D.
This suggests that the two lightcone coordinates $u$ and $w$ should be identified as the coordinates of the 2D space.
The 4D d'Alembertian in lightcone coordinates, $\Box = 4 (\partial_u \partial_v - \partial_w \partial_{\bar{w}})$, cannot be interpreted as the 2D equivalent, so the kinetic term of SDYM must be scrapped and replaced with the correct 2D version.
This argument should be understood purely as an alternative route to ZM theory.
While NLSM and ZM are the same theory classically when restricted to 2D, SDYM and ZM are not because, for example, the tree amplitudes of SDYM vanish but those of ZM do not (cite argumentative papers).

@@

\subsubsection{Wess-Zumino-Witten theory -- James}
I want to write this down or I'm going to completely forget it

@@

\subsubsection{Cubic 2-form theory -- James}

@@

\subsubsection{Chern-Simons theory -- James}
We won't use this but it's useful to treat all of the CK theories in one place
Cite Henrik Maor
Again you place some constraint on the field strength and use Lorenz gauge

\subsubsection{Generalities of known color-dual theories -- James}
Should this section maybe go first?

>>>>

\paragraph{Jacobi relations}
\label{sec:jac-rels}

@@

\paragraph{Graph symmetries}
\label{sec:syms}

@@

\paragraph{Unitarity}
\label{sec:unitarity}

@@

\section{One-loop cubic construction}
Here we'll describe the why constructing the full Yang-Mills kinematic algebra (outside self-dual sector) is generically hard, and why the same reason makes NLSM beyond one-loop hard as well. It has to do with out the vector state sum mixes with the four-point contact. In terms of our previously defined Feynman rules, the three-point YM vertex takes the form:
\be
\cubic{hgrey0}{gluon}{}{gluon}{}{gluon}{}  =  \cubic{hgrey0}{fermion2}{}{}{nhpRed}{}{nhpRed} + \cubic{hgrey0}{}{nhpRed}{fermion2}{}{}{nhpRed} + \Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{} +\text{h.c.}
\ee
The state-sum between two three-point gluon vertex yields the following expansion in terms of YZ-theory vertices:
\be
 \AcubicB{hgrey0}{gluon}{}{gluon}{}{gluon}{}{gluon}{} =  \AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{fermion2}{} + 5 \text{ perms} + \AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{}{nhpRed}
\ee
All the $(\epsilon\epsilon)^1$ diagram permutations are isolated to the cubic sector and together manifest color-kinematics duality. However, the exception of the last diagram which introduces a $(\epsilon\epsilon)^2$ factor, must conspire with the Yang-Mills contact to preserve the duality off-shell

\be
\AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{}{nhpRed} = (s_{13}-s_{12}) \epsilon_{(12)} \epsilon_{(34)}
\ee
One was to resolve this is to introduce higher spin modes in order to accomodate the internal index contractions needed to recover the s-channal contract:
\be
\AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{Bfield}{nhpRed}{}{nhpRed} = s_{12 }(\epsilon_{(13)} \epsilon_{(24)} - \epsilon_{(14)} \epsilon_{(23)})
\ee
While introducing succesively higher spiin states to the NMHV lagrangian constructed in \citepls{} could in principle work at tree level, we will see that it is not consiisntent with what we find at general loop order. We will comment on this in \sect{}. For now, we will study how to construct one-loop integrands consistent with the color-kinematics by isolating the cubic sector of the state sum above. This is easily understood in terms of a polarization decomposition introduced by one of the authors \citepls{} that decomposes vector amplitudes in terms of building blocks that preserve partial amplitude relations. 
\subsection{Block Decomposition of Yang-Mills}
Describe how different orders in $\varepsilon_i\varepsilon_j$ in pure Yang-Mills captures different cubic and contact sectors of the theory. All the $D$-dimensional Lagnrangians above can be understood as selecting specific on shell states to isolate the cubic sector of Yang-Mills at a given loop order. Start with reducible amplitude block decomp from Ref. \citepls{}:
\begin{eBox}
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
\end{eBox}
By a simple mass-dimensioin argument, we know the tree-level cubic sector is simply:
\be
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
and likewise at one-loop:
\be
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
We can select out both of these manifestly cubic sectors by selecting only $\epsilon^+\epsilon^+$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level OR what $J$-theory and $YZ$-theory select with their on-shell states. Indeed, the $D$-dimensional description of $J$ theory is just the following blocks in the RABD expansion of Yang-Mills:
\be
A(...,\bar{J}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
\ee
and similarly so at one-loop:
\be
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}
\ee
To recover NLSM at one-loop, we need to extract out the $D$-dependent part that corresponds to an internal $YY$-loop from extra-dimensional scalars:
\begin{eBox}
\be
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\epsilon\rightarrow k}
\ee
\end{eBox}
Why must we take the derivative with respect to $D$? After all, $J$-theory and $YZ$-theory both have well defined propagators for internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we noted in the previous section, all the $J$ and $Z$ states must be on-shell in order to produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at one-loop will not produce NLSM amplitudes. However, the internal $YY$-loop is a valid forward limit, since as constructed YZ theory is valid for off-shell Y-particles, and thus they serve as a valid forward limit. We will construct these one-loop integrands below in the next section. 
\subsection{Semi-abelian Yang-Mills theory}
\subsection{One-loop SDYM integrands}
\be
\simplebox = g_{\text{YM}}^4 X(k_1,l_1)X(k_2,l_2)X(k_3,l_3)X(k_4,l_4)
\ee
\subsection{One-loop NLSM Integrands}

Note from James:  I've explicitly checked the nice form of the 1-loop numerators at 4pt, 6pt, and 8pt.
\be
\simplebox = (2 f_\pi^{-1})^4 (l_1 k_1)(l_2 k_2)(l_3 k_3)(l_4 k_4)
\ee
where we define sequential loop momenta on the $n$-gon as $l_{i+1}= l_i+k_i$. Given this, the $n$-gon permits a equivalent expression in terms of inverse propagators:
\be
\simplebox =f_\pi^{-4} [12][23][34][41]
\ee
Where we have define the antisymmetric kinematic variable, $[LR] = l_i^L - l_R^2$
\subsection{Two-loop obstruction}

\section{Two-loop bootstrap}

@@

\subsection{Double-copy verification}
\ace{I'm punting a lot of discussion to Ref \cite{Carrasco:2023qgz}.
  Do we think that is a problem?}

\begin{figure}
  \begin{center}
	\PhysicalCutOne{}{}{}{} \PhysicalCutTwo{}{}{}{}
  \end{center}
  \caption{The two physical cut topologies shared by sGal, BI, and DBIVA}
  \label{fig:emu}
\end{figure}

@@

\subsection{Two-loop Yang-Mills revisited}
Given the close ties between pion and gluon scattering for trees and
at one loop, the success of a globally color-dual solution for pions
at two loops warrants a reexamination of the two-loop pure Yang--Mills
problem.  As discussed above in \cref{sec:bootstrap}, BDN failed at
finding an off-shell color-dual representation satisfying 7
assumptions, and were forced to develop an on-shell notion of
color-kinematics duality \cite{Bern:2015ooa}.  Thus we must decide
which assumption(s) to break.  From both a motivation and
implementation standpoint, the simplest assumption to break is the
``power-counting'' statement, which we analyze first.

@@

      \ver{c1}{0:\cl};
      \ver{c2}{\cang:\cl};
      \ver{c3}{2*\cang:\cl};
      \ver{c4}{3*\cang:\cl};
      \ver{ci}{$(i5)!0.5!(i1)$};
      \path let \p1 =(ci) in coordinate (c5) at (\x1,0.2){};
      \draw[ultra thick] (e1)--(i1);
      \draw[ultra thick] (e2)--(i2);
      \draw[ultra thick] (e3)--(i3);
      \draw[ultra thick] (e4)--(i4);
      \draw[ultra thick] (i5)--(i1)--(i4)--(i3)--(i2)--(i6);
      \draw[ultra thick] (i6) to[out=60,in=180-60] (i5);
      \draw[ultra thick] (i6) to[out=-60,in=180+60] (i5);
      \draw[dashed,ultra thick,nhp4] (0,0) -- (c1);
      \draw[dashed,ultra thick,nhp4] (0,0) -- (c2);
      \draw[dashed,ultra thick,nhp4] (0,0) -- (c3);
      \draw[dashed,ultra thick,nhp4] (0,0) -- (c4);
      \draw[dashed,ultra thick,nhp4] (c5) -- ++(90:0.7*\cl);
      \node[fill=white,circle] (c) at (0,0) {};

@@

\ace{\sout{I need to check this.  For now it is a standing
    hypothesis:} Checked.  The failure works as stated.} In fact, we can make the
failure extremely precise.  Start with the diagrams shown in the
Jacobi relation from \cref{eq:TwoLoopJacobiOne}: the double-box,
crossed-box, and penta-triangle.  For each of the three diagrams,
write down the most generic parity-even local ansatz involving four
polarizations and six momenta (any combination of external or loop).
Each ansatz will be 10,010 terms initially.  Next impose the symmetry
constraints on each of the three diagrams, leaving 2761 free
parameters on the double-box, 2576 free parameters on the crossed-box,
and 5040 free paramters on the penta-triangle.  Then impose the
``kissing triangle'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only recieves contributions from the
double-box
\begin{equation}
  \KissingTriangles
  =
  \frac{1}{(\textcolor{nhpBlue}{\ell_1})^2} \
  n \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{(\textcolor{nhpBlue}{\ell_2})^2} \
  n\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied:
\begin{equation}
  \KissingTriangles
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang--Mills representation: locality,
symmetry, and the ``kissing triangle'' cut are incompatible with the
kinematic Jacobi relations.

@@

\section{All-loop no-go theorem}
This could be a section about our expectations for all-loop construction requiring non-local color-dual integrands. Lay out counterfactual proof, i,e. define the following statement:
\be
 Q_L=\textit{``the $L$-loop numerator is polynomial and color-dual"}. 
\ee
Then it is clear that if the kinematic numerator is polynomial and color-dual, evaluating it in kinematic regions where the internal legs are on-shell won't disrupt the kinematic algrebraic relations, and it must still be polynomial. Thus we find:
\be
Q_L \Rightarrow Q_{L-1}
\ee
Or equivalently,
\be
\text{not }Q_{L-1}\Rightarrow \text{not }Q_L
\ee
We have found that the two-loop integrand for Yang-Mills cannot be constructed from a color-dual polynomial kinematic numerator. Thus, if a color-dual representation exists at all-loop order for Yang-Mills it must be a rational function of kinematics.

@@

\section{Catalog of color-dual theories}

@@

\paragraph{Self-Dual Yang-Mills} The first example in the literature \citepls{} of color-dual Feynman rules is that of Yang-Mills theory in the self-dual sector (SDYM):

@@

\paragraph{YZ-Theory}
SDYM makes explicit reference to the spacetime dimension and Zakharov-Mikhailov theory, to be discussed in \cref{sec:ZMTheory}, does as well.
This makes both theories poorly suited for constructing dimensionally-regulated loop-level amplitudes.
On the other hand, YZ-theory is an honest $D$-dimensional theory that captures the classical physics of the nonlinear sigma model (NLSM) while manifest the duality between color and kinematics \cite{Cheung:2016prv}.

@@

\paragraph{J-theory} There is an alternative color-dual formulation of NLSM that the theory manifestly corresponds to NLSM at least at tree level \cite{Cheung:2021zvb}.
For related formulations of NLSM that do not treat color-kinematics duality see \cite{Freedman:1980us, Slavnov:1971mz}.
The construction begins with a first order formulation of NLSM in terms of chiral currents $J^\mu$.
First order formulation of NLSM in terms of chiral current $J^\mu$.
Two equations.
Two constraints, the field strength should vanish
\be
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0
\ee
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge
\be
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0
\ee
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\be
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\ee
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian
\begin{eBox}
\be
\mathcal{L}^{\text{J}} =\bar{J}^\mu \Box J_\mu +g f^{abc} \bar{J}^\mu J^\nu \partial_\nu J^\mu .
\ee
\end{eBox}
By calculating an off-shell 4pt correlation function it is possible to check that this Lagrangian automatically satisfies color-kinematics duality.
Kinematic algebra is diff algebra (as usual).
How to extract physical states... (longitudinal J, weird J bar, restores broken Bose).

@@

\paragraph{Zakharov-Mikhailov theory}
\label{sec:ZMTheory}
While J-theory manifests color-kinematics duality, it only does so at tree level because J-theory breaks manifest Bose symmetry.
Manifest color-kinematics duality can be achieved at all loop orders by restricting J-theory to two spacetime dimensions where the chiral current can be dualized to a scalar $J^\mu = \epsilon^{\mu\nu} \partial_\nu \varphi$, which trivially enforces Lorenz gauge.
After plugging $\varphi$ into \cref{eq:JTheoryEOM} and rearranging, another copy of $\varphi$ (rather than some new field) can be integrated in to obtain the Zakharov-Mikhailov (ZM) Lagrangian
\begin{eBox}
\be
\label{eq:ZMLagrangian}
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2 + g f^{abc} \varphi^a \epsilon^{\mu\nu}(\partial_\mu \varphi^b)( \partial_\nu \varphi^c) =  \text{tr}\left[\frac{1}{2}(\partial \varphi)^2-i g \varphi [\partial_t \varphi, \partial_z \varphi ]\right] .
\ee
\end{eBox}
This procedure makes it clear that ZM and NLSM encode the same classical physics, that is, their equations of motion are dual to each other.
ZM theory can also be tenuously motivated from SDYM theory.
The power counting of the interaction term in the SDYM equation of motion \cref{SDEOM} makes it clear that the theory is critical in 2D.
This suggests that the two lightcone coordinates $u$ and $w$ should be identified as the coordinates of the 2D space.
The 4D d'Alembertian in lightcone coordinates, $\Box = 4 (\partial_u \partial_v - \partial_w \partial_{\bar{w}})$, cannot be interpreted as the 2D equivalent, so the kinetic term of SDYM must be scrapped and replaced with the correct 2D version.
This argument should be understood purely as an alternative route to ZM theory.
While NLSM and ZM are the same theory classically when restricted to 2D, SDYM and ZM are not because, for example, the tree amplitudes of SDYM vanish but those of ZM do not (cite argumentative papers).

@@

\paragraph{Wess-Zumino-Witten theory -- James}
I want to write this down or I'm going to completely forget it

@@

\paragraph{Cubic 2-form theory -- James}

@@

\paragraph{Chern-Simons theory -- James}
We won't use this but it's useful to treat all of the CK theories in one place
Cite Henrik Maor
Again you place some constraint on the field strength and use Lorenz gauge

\paragraph{Generalities of known color-dual theories -- James}
Should this section maybe go first?

\begin{itemize}
\item There's also CS but the amplitudes are all zero (and marginally 10D N=1 SYM)
\item All of the theories described above have diff as kinematic algebra, SU(N) for BAS, or rigid/spin Lorenz for cubic 2-form theory
\item All of the theories choose a simple gauge like Lorenz or lightcone
\item All of the theories place some constraint on the field strength directly and get their equations of motion directly from that constraint, not the YM EOM.
\item (Almost all of the theories (aside from NLSM) have trivial amplitudes at least at tree level)
\item This means that almost all of the theories only have CK at the classical level, which can sometimes be fudged into 1-loop as we'll now discuss.
\item For off-shell CK, the kinetic term is relatively unimportant so long as the poles are correctly/consistently identified (what do you mean by s,t,u channel poles).  This might screw up locality in the squared theory.
\end{itemize}
\section{Spinor-helicity and conventions}
